### 2015-03-12 Paper/content/00_abstract.tex
paper: shiny new structure ^-^

\begin{abstract}
...
\end{abstract}

\begin{IEEEkeywords}
cloud computing, deployment models, delivery models
\end{IEEEkeywords}