### 2015-03-12 Paper/content/06_facts.tex

\section{Stand der Technik}
\label{sec_facts}

\IEEEPARstart{S}{tand}...