### 2015-03-12 Paper/content/07_related_work.tex

\section{Verwandte Arbeiten}
\label{sec_related_work}

\IEEEPARstart{V}{erwandte} Arbeiten...