### 2015-03-12 Paper/content/08_conclusion.tex
responsibilities for sections :)

\section{Ergebnisse und Ausblick}
\label{sec_conclusion}

\IEEEPARstart{Z}{usammenfassung}...