### 2015-03-22 Paper/content/00_abstract.tex

\begin{abstract}
...
\end{abstract}

\begin{IEEEkeywords}
cloud computing, comparison, deployment models, delivery models
\end{IEEEkeywords}