### 2015-03-22 Paper/content/10_references.tex
references added

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

\begin{thebibliography}{1}

\bibitem{vortragSAP}
Martin Heisig. "Public Cloud vs. Private Cloud: Cloud Computing - Wie die Industrie sich neuen Herausvorderungen stellt", Vortrag am Hasso-Plattner-Institut, Potsdam, 27.11.2014.

\bibitem{vortragFujitsu}
Karsten Beins. "Flexible options to make Cloud a viable choice for enterprise IT", Vortrag am Hasso-Plattner-Institut, Potsdam, 04.12.2014.

\bibitem{binnig2009}
Carsten Binnig, Donald Kossmann, Tim Kraska und Simon Loesing. "How is the Weather tomorrow?: Towards a Benchmark for the Cloud", Proceedings of the Second International Workshop on Testing Database Systems (DBTest '09), ACM, New York, NY, USA, 2009.

\bibitem{li2010}
Ang Li, Xiaowei Yang, Srikanth Kandula und Ming Zhang. "CloudCmp: Comparing Public Cloud Providers", Proceedings of the 10th ACM SIGCOMM conference on Internet measurement, ACM, 2010.

\bibitem{garg2011}
Saurabh Kumar Garg, Steven Versteeg und Rajkumar Buyya. "SMICloud: A Framework for Comparing and Ranking Cloud Services", Fourth IEEE International Conference on Utility and Cloud Computing (UCC), IEEE, 2011.

\bibitem{zhang2013}
Xiao Zhang, Feng WenXiong und Xiao Qin. "Performance Evaluation of Online Backup Cloud Storage", International Journal of Cloud Applications and Computing (IJCAC) 3.3 (20-33), 2013.

\bibitem{zenuni2014}
Xhemal Zenuni, Jaumin Ajdari, Florije Ismaili und Bujar Raufi. "Cloud Storage Providers: A Comparison Review and Evaluation", Proceedings of the 15th International Conference on Computer Systems and Technologies, ACM, 2014.

\bibitem{ibm20011}
Dan Orlando. "Cloud computing service models", IBM developerWorks, 2011.

\bibitem{nistDefinition}
Peter Mell, Timothy Grance. "The NIST Definition of Cloud Computing", NIST Special Publication 800-145, 2011.

\bibitem{rackSpace}
Ben Kepes. "Understanding The Cloud Computing Stack", CloudU, 2011.

\bibitem{armbrust}
Michael Armbrust, Armando Fox, Rean Griffith, Anthony D. Joseph, Randy Katz, Andy Konwinski, Gunho Lee, Dav id Patterson, Ariel Rabkin, Ion Stoica und Matei Zaharia. "A View of Cloud Computing", Communications of the ACM Vol. 53 No. 4 (50-58), 2010.

\bibitem{awsOverview}
Jinesh Varia, Sajee Mathew. "Overview of Amazon Web Services", Amazon, 2014.

\bibitem{nistStandards}
Michael Hogan, Annie Sokol. "NIST Cloud Computing Standards Roadmap", NIST Special Publication 500-291 Version 2, 2011.

\bibitem{technet}
Bill Loeffler. "Cloud Computing: What is Infrastructure as a Service", 22.03.2015, \url{https://technet.microsoft.com/en-us/magazine/hh509051.aspx}.

\bibitem{privateFail}
Matt Asay. "Private cloud's very public failure", 22.03.2015, \url{http://www.techrepublic.com/article/private-clouds-very-public-failure/}.

\bibitem{openGroup}
The Open Group. "Cloud Portability and Interoperability", 22.03.2015, \url{http://www.opengroup.org/cloud/cloud/cloud_iop/cloud_port.htm}.

\bibitem{gigaOm}
Tim Crawford. "The enterprise view of cloud, specifically Private Cloud, is confusing", 22.03.2015, \url{http://research.gigaom.com/2015/02/the-enterprise-view-of-cloud-specifically-private-cloud-is-confusing/}.

\bibitem{dbWsj}
Steven Norton. "Deutsche Bank: CIOs Continue Move to Private Cloud", 22.03.2015, \url{http://blogs.wsj.com/cio/2015/01/09/deutsche-bank-cios-continue-move-to-private-cloud/}.

\bibitem{dynamo}
Werner Vogels. "Amazon DynamoDB – a Fast and Scalable NoSQL Database Service Designed for Internet Scale Applications", 22.03.2015, \url{http://www.allthingsdistributed.com/2012/01/amazon-dynamodb.html}.

\bibitem{openStack}
The OpenStack Project. "OpenStack: The Open Source Cloud Operating System", 22.03.2015, \url{http://www.openstack.org/software/}.

\bibitem{dockerZdnet}
Steven J. Vaughan-Nichols. "What is Docker and why is it so darn popular?", 22.03.2015, \url{http://www.zdnet.com/article/what-is-docker-and-why-is-it-so-darn-popular/}.

\bibitem{hybridGartner}
Tom Bittman. "Hybrid Cloud is Three Years Behind Private Cloud", 22.03.2015, \url{http://blogs.gartner.com/thomas_bittman/2013/10/24/hybrid-cloud-is-three-years-behind-private-cloud/}.

\bibitem{privateFail2}
Tom Bittman. "Why Are Private Clouds Failing", 22.03.2015, \url{http://blogs.gartner.com/thomas_bittman/2014/09/12/why-are-private-clouds-failing/}.

\bibitem{amazonPrivate}
Brandon Butler. "Would Amazon Web Services ever build an on-premise private cloud", 22.03.2015, \url{http://www.networkworld.com/article/2851328/cloud-computing/would-amazon-web-services-ever-build-an-on-premise-private-cloud.html}.

\end{thebibliography}