### 2015-03-22 Paper/content/10_references.tex

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

\shorthandoff{"}

\begin{thebibliography}{1}

\bibitem{nistStandards}
Michael Hogan, Annie Sokol. "NIST Cloud Computing Standards Roadmap", NIST Special Publication 500-291 Version 2, 2011.

\bibitem{rackSpace}
Ben Kepes. "Understanding The Cloud Computing Stack", CloudU, 2011.

\bibitem{technet}
Bill Loeffler. "Cloud Computing: What is Infrastructure as a Service", 22.03.2015, \url{https://technet.microsoft.com/en-us/magazine/hh509051.aspx}.

\bibitem{ibm2011}
Dan Orlando. "Cloud computing service models", IBM developerWorks, 2011.

\bibitem{vortragFujitsu}
Karsten Beins. "Flexible options to make Cloud a viable choice for enterprise IT", Vortrag am Hasso-Plattner-Institut, Potsdam, 04.12.2014.

\bibitem{fujitsuIaaS}
"FUJITSU Cloud Infrastructure as a Service". http://www.fujitsu.com/global/services/infrastructure/iaas/, Zugriff 22.03.2015

\bibitem{fujitsuHybrid}
"Hybrid Cloud Services". http://www.fujitsu.com/global/services/application-services/application-development-Integration/hybridcloud/index.html, Zugriff 22.03.2015

\bibitem{fujitsuPaaS}
"FUJITSU Cloud Platform as a Service". http://www.fujitsu.com/global/services/application-services/paas/index.html, Zugriff 22.03.2015

\bibitem{fujitsuSaaS}
"Fujitsu Software-as-a-Service". http://www.fujitsu.com/de/cloud/solutions/saas/, Zugriff 22.03.2015

\bibitem{amazonPrivate}
Brandon Butler. "Would Amazon Web Services ever build an on-premise private cloud", 22.03.2015, \url{http://www.networkworld.com/article/2851328/cloud-computing/would-amazon-web-services-ever-build-an-on-premise-private-cloud.html}.

\bibitem{privateFail}
Matt Asay. "Private cloud's very public failure", 22.03.2015, \url{http://www.techrepublic.com/article/private-clouds-very-public-failure/}.

\bibitem{openGroup}
The Open Group. "Cloud Portability and Interoperability", 22.03.2015, \url{http://www.opengroup.org/cloud/cloud/cloud_iop/cloud_port.htm}.

\bibitem{gigaOm}
Tim Crawford. "The enterprise view of cloud, specifically Private Cloud, is confusing", 22.03.2015, \url{http://research.gigaom.com/2015/02/the-enterprise-view-of-cloud-specifically-private-cloud-is-confusing/}.

\bibitem{dynamo}
Werner Vogels. "Amazon DynamoDB – a Fast and Scalable NoSQL Database Service Designed for Internet Scale Applications", 22.03.2015, \url{http://www.allthingsdistributed.com/2012/01/amazon-dynamodb.html}.

\bibitem{openStack}
The OpenStack Project. "OpenStack: The Open Source Cloud Operating System", 22.03.2015, \url{http://www.openstack.org/software/}.

\bibitem{dockerZdnet}
Steven J. Vaughan-Nichols. "What is Docker and why is it so darn popular?", 22.03.2015, \url{http://www.zdnet.com/article/what-is-docker-and-why-is-it-so-darn-popular/}.

\bibitem{privateFail2}
Tom Bittman. "Why Are Private Clouds Failing", 22.03.2015, \url{http://blogs.gartner.com/thomas_bittman/2014/09/12/why-are-private-clouds-failing/}.

\bibitem{binnig2009}
Carsten Binnig, Donald Kossmann, Tim Kraska und Simon Loesing. "How is the Weather tomorrow?: Towards a Benchmark for the Cloud", Proceedings of the Second International Workshop on Testing Database Systems (DBTest '09), ACM, New York, NY, USA, 2009.

\bibitem{li2010}
Ang Li, Xiaowei Yang, Srikanth Kandula und Ming Zhang. "CloudCmp: Comparing Public Cloud Providers", Proceedings of the 10th ACM SIGCOMM conference on Internet measurement, ACM, 2010.

\bibitem{garg2011}
Saurabh Kumar Garg, Steven Versteeg und Rajkumar Buyya. "SMICloud: A Framework for Comparing and Ranking Cloud Services", Fourth IEEE International Conference on Utility and Cloud Computing (UCC), IEEE, 2011.

\bibitem{zhang2013}
Xiao Zhang, Feng WenXiong und Xiao Qin. "Performance Evaluation of Online Backup Cloud Storage", International Journal of Cloud Applications and Computing (IJCAC) 3.3 (20-33), 2013.

\bibitem{zenuni2014}
Xhemal Zenuni, Jaumin Ajdari, Florije Ismaili und Bujar Raufi. "Cloud Storage Providers: A Comparison Review and Evaluation", Proceedings of the 15th International Conference on Computer Systems and Technologies, ACM, 2014.

\bibitem{toosi2014}
Adel Nadjaran Toosi, Rodrigo N. Calheiros und Rajkumar Buyya. "Interconnected Cloud Computing Environments: Challenges, Taxonomy and Survey", ACM Computing Surveys (CSUR) 47.1, 2014.

\end{thebibliography}